On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-11-29 manuscripts/PNAS/second review/resub/chazthm_pnas_resub.tex
end of proof of lemma 1 explanation

\begin{proof}[Proof of Thm.~\ref{DeterministicUniquenessCorollary} for $k=1$]
Since the only 1-uniform hypergraph with the SIP is $[m]$, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case, we may take any $C_1 \geq 1/ \min_{\ell \in [m]} |c_{\ell}|$.

@@

We next show that \eqref{cond} holds. Fixing $\mathcal{G} \subseteq {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1}$, it suffices to show that $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}} ) < 1$, since by \eqref{dimLem} we then have $|\cap \pi(\mathcal{G})| = \dim(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}) \leq \dim(\bm{\mathcal{A}}_{\cap \mathcal{G}}) = |\cap \mathcal{G}|$, with equalities by the full column-ranks of $\mathbf{A}_{S}$ and $\mathbf{B}_{\pi(S)}$ for all $S \in \mathcal{H}$.\footnote{Note that if ever $\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})} \neq \bf 0$ while $\cap \mathcal{G} = \emptyset$, then we would have $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm 0 ) = 1$. This cannot be the case, however, since the deduction that follows would then lead to a contradiction.} Observe that $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}}  ) 
\leq d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \cap \bm{\mathcal{A}}_\mathcal{G} \right)$, since $\bm{\mathcal{A}}_{\cap \mathcal{G}} = \cap \bm{\mathcal{A}}_\mathcal{G}$ by Lem.~\ref{spanIntersectionLemma} and $d(U',V) \leq d(U,V)$ whenever $U' \subseteq U$. Recalling Def.~\ref{dDef}, we may apply Lem.~\ref{DistanceToIntersectionLemma} and carry the supremum through the sum:
\begin{align}
d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \cap \bm{\mathcal{A}}_\mathcal{G} \right)
%&\leq \sup_{\mathbf{u} \in \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \ \|\mathbf{u}\|_2 \leq 1} \sum_{S \in \mathcal{G}} \frac{ \text{\rm dist}\left( \mathbf{u},\bm{\mathcal{A}}_{S} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) }
\leq \sum_{S \in \mathcal{G}} \frac{ d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})},\bm{\mathcal{A}}_{S} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) }.
\end{align}
Since $\cap \bm{\mathcal{B}}_{\pi(\mathcal{G})} \subseteq \bm{\mathcal{B}}_{\pi(S)}$ for all $S \in \mathcal{G}$, the numerator of each term above is bounded by $d\left( \bm{\mathcal{B}}_{\pi(S)},\bm{\mathcal{A}}_{S} \right)$ and, recalling \eqref{eqdim} (since $\dim(\bm{\mathcal{B}}_{\pi(S)}) = \dim(\bm{\mathcal{A}}_S)$), by $d\left(\bm{\mathcal{A}}_{S}, \bm{\mathcal{B}}_{\pi(S)} \right) \leq \varepsilon$, in turn. Thus,
\begin{align}
d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}} )
\leq \frac{|\mathcal{G}| \varepsilon}{1 - \xi( \bm{\mathcal{A}}_\mathcal{G} )}
\leq \frac{C_2 \varepsilon}{\max_j\|\mathbf{A}_j\|_2},
\end{align}
which is less than one, since $C_2 \varepsilon < L_2(\mathbf{A}) \leq \max_j\|\mathbf{A}_j\|_2$.

@@

By Lem.~\ref{NonEmptyLemma}, the association $i \mapsto \cap \pi(\sigma(i))$ is an injective map $\overline \pi: J \to [\overline m]$ for some $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$, and $\mathbf{B}_{\overline \pi(i)} \neq \mathbf{0}$ for all $i \in J$ since the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent for all $S \in \mathcal{H}$. It then follows from \eqref{eqdim} and \eqref{randoml} that $d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \leq C_2 \varepsilon / \max_i \|\mathbf{A}_i\|_2$ for all $i \in J$. Fixing $\overline \varepsilon = C_2\varepsilon$ and letting $c_i = \|\mathbf{A}_i\|_2^{-1}$, we thus have that for each $\mathbf{e}_i \in \mathbb{R}^m$ with $i \in J$ there exists some $\overline{c}_i \in \mathbb{R}$ such that $\|c_i\mathbf{A}\mathbf{e}_i - \overline{c}_i \mathbf{B}\mathbf{e}_{\overline \pi(i)}\|_2 \leq \overline \varepsilon < L_2(\mathbf{A}) \min_{i\in J} |c_i|$.  But this is exactly the supposition in \eqref{1D}, and the result follows from the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$.
\end{proof}

>>>>

\begin{proof}[Proof of Thm.~\ref{DeterministicUniquenessCorollary} for $k=1$]
Since the only 1-uniform hypergraph with the SIP is $[m]$, we have $\mathbf{x}_i = c_i \mathbf{e}_i$ for $c_i \in \mathbb{R} \setminus \{\mathbf{0}\}$, $i \in [m]$. In this case, we may take $C_1 = 1/ \min_{\ell \in [m]} |c_{\ell}|$.

@@

We next show that \eqref{cond} holds. Fixing $\mathcal{G} \subseteq {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1}$, it suffices to show that $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}} ) < 1$, since by \eqref{dimLem} we then have $|\cap \pi(\mathcal{G})| = \dim(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}) \leq \dim(\bm{\mathcal{A}}_{\cap \mathcal{G}}) = |\cap \mathcal{G}|$, with equalities by the full column-ranks of $\mathbf{A}_{S}$ and $\mathbf{B}_{\pi(S)}$ for all $S \in \mathcal{H}$.\footnote{Note that if ever $\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})} \neq \bf 0$ while $\cap \mathcal{G} = \emptyset$, then we would have $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm 0 ) = 1$. This cannot be the case, however, since the deduction that follows would then lead to a contradiction.} Observe that $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}}  ) 
\leq d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \cap \bm{\mathcal{A}}_\mathcal{G} \right)$, since $\bm{\mathcal{A}}_{\cap \mathcal{G}} = \cap \bm{\mathcal{A}}_\mathcal{G}$ by Lem.~\ref{spanIntersectionLemma} and $d(U',V) \leq d(U,V)$ whenever $U' \subseteq U$. Recalling Def.~\ref{dDef}, we may apply Lem.~\ref{DistanceToIntersectionLemma} and carry the supremum through the sum:
\begin{align}
d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \cap \bm{\mathcal{A}}_\mathcal{G} \right)
%&\leq \sup_{\mathbf{u} \in \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \ \|\mathbf{u}\|_2 \leq 1} \sum_{S \in \mathcal{G}} \frac{ \text{\rm dist}\left( \mathbf{u},\bm{\mathcal{A}}_{S} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) }
\leq \sum_{S \in \mathcal{G}} \frac{ d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})},\bm{\mathcal{A}}_{S} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) }.
\end{align}
Since $\cap \bm{\mathcal{B}}_{\pi(\mathcal{G})} \subseteq \bm{\mathcal{B}}_{\pi(S)}$ for all $S \in \mathcal{G}$, the numerator of each term above is bounded by $d\left( \bm{\mathcal{B}}_{\pi(S)},\bm{\mathcal{A}}_{S} \right)$ and, recalling \eqref{eqdim} (since $\dim(\bm{\mathcal{B}}_{\pi(S)}) = \dim(\bm{\mathcal{A}}_S)$), by $d\left(\bm{\mathcal{A}}_{S}, \bm{\mathcal{B}}_{\pi(S)} \right) \leq \varepsilon$, in turn. Thus,
\begin{align}\label{last}
d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}} )
\leq \frac{|\mathcal{G}| \varepsilon}{1 - \xi( \bm{\mathcal{A}}_\mathcal{G} )}
\leq \frac{C_2 \varepsilon}{\max_j\|\mathbf{A}_j\|_2},
\end{align}
which is less than one, since $C_2 \varepsilon < L_2(\mathbf{A}) \leq \max_j\|\mathbf{A}_j\|_2$.

@@

By Lem.~\ref{NonEmptyLemma}, the association $i \mapsto \cap \pi(\sigma(i))$ is an injective map $\overline \pi: J \to [\overline m]$ for some $J \subseteq [m]$ of size $\overline m - r(\overline m - m)$, and $\mathbf{B}_{\overline \pi(i)} \neq \mathbf{0}$ for all $i \in J$ since the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent for all $S \in \mathcal{H}$. It then follows from \eqref{eqdim} and \eqref{last} that $d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \leq \overline \varepsilon$ for all $i \in J$, with $\overline \varepsilon := C_2 \varepsilon / \max_i \|\mathbf{A}_i\|_2$. %Fixing $\overline \varepsilon = C_2\varepsilon$ and letting $c_i = \|\mathbf{A}_i\|_2^{-1}$, we thus have that for each $\mathbf{e}_i \in \mathbb{R}^m$ with $i \in J$ there exists some $\overline{c}_i \in \mathbb{R}$ such that $\|c_i\mathbf{A}\mathbf{e}_i - \overline{c}_i \mathbf{B}\mathbf{e}_{\overline \pi(i)}\|_2 \leq \overline \varepsilon < L_2(\mathbf{A}) \min_{i\in J} |c_i|$. But this is exactly the supposition in \eqref{1D}, and the result follows from the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$. 
Letting $c_i := \|\mathbf{A}_i\|_2^{-1}$, we have by Def.~\ref{dDef}:
\begin{align*}
d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right)
&= \sup_{\bm{u} \in \bm{\mathcal{A}}_i, \|\bm{u}\|_2 \leq 1} \text{dist}\left(\bm{u}, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \\
&\geq  \text{dist}\left(c_i\bm{Ae}_i, \mathbf{Be}_{\overline \pi(i)} \right) \\
&= \inf_{\overline x_i \in \mathbb{R}} \|c_i\bm{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2
\end{align*}
%
and there must then exist some $\overline c_i$ such that $\|c_i\bm{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2 \leq \overline \varepsilon$ for $\overline \varepsilon < L_2(\mathbf{A}) \min_{i\in J} |c_i|$ (since $\varepsilon < L_2(\mathbf{A})/C_2$). But this is exactly the supposition in \eqref{1D}, and the result follows from the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$. 
\end{proof}
